---
title: "PSTAT 174 PROJECT"
author: "Danny Baerman, Vela Hou, Lilian Li, Janelle Samansky, Andy Tran"
date: "February 21, 2019"
output: pdf_document
---
Libraries
```{r}
#Load Packages
library(MASS)
library(forecast)
library(tseries)
library(astsa)
library(dse)
library(knitr)
library(gridExtra)
library(grid)

```


Read in CSV, Initial Time Series Analaysis
```{r}

beer <- read.csv(file = "beer.csv",  header = TRUE, sep = ",", nrows = 146)[,-1]

beerall <- read.csv(file = "beer.csv",  header = TRUE, sep = ",", nrows = 154)[,-1]

ts.plot(beer, ylab = "Production in Megalitres",xlab = "Time(Quarterly)")
title(expression(Beer~Production~In~Australia~March~1956~-~June~1992))


```

ACF and PACFs for Beer Production
```{r}
# Beer.ts with 8 points removed
beer.ts <- ts(beer, frequency = 1)

# Beer.ts with all points 
beerall.ts <- ts(beerall, frequency = 1)

acf(beer.ts, lag.max =10,main = "" )
pacf(beer.ts, lag.max=10, main = "")
```

Seasonal Plot
```{r}
# Seasonal Plot
seasonplot(beer.ts, 4, col=rainbow(3), year.labels = TRUE, main = "Seasonal~Plot")

```

Box Cox Transformations
```{r}
# Finding Lambda
bcbeer <- boxcox(beer.ts~as.numeric(1:length(beer)))
# Lambda
lambda1 <- bcbeer$x[which.max(bcbeer$y)]
# Lambda is 1/2
# Log transformed data
beer.tr <- beer.ts^(1/2)
ts.plot(beer.ts, ylab = "Production in Megalitres",xlab = "Time(Quarterly)", main = "Box Cox Transformed Data")

# Variance of nontransformed data
var(beer.ts)
# Variance of transformed data
var(beer.tr)
```


Removing Trend and Seasonality
```{r}
# Differencing Seasonality at lag 4
beerdiff4 <- diff(beer.tr, lag =4)
var(beerdiff4)
ts.plot(beerdiff4, ylab = "Differenced At Lag 4 ")
abline(lm(beerdiff4~as.numeric(1:length( beerdiff4))), col ="red")

# Differencing at Lag 1
beerdiff4diff1 <- diff(beerdiff4, lag =1)
var(beerdiff4diff1)
ts.plot(beerdiff4diff1, ylab = "Differenced At Lag 4, Lag 1")
abline(lm(beerdiff4diff1~as.numeric(1:length( beerdiff4diff1))), col ="red")


```


```{r}
# Differenced at lag 4 PACF AND ACF
par(mfrow=c(1,2))
acf(beerdiff4, lag.max = 24, main = "")
pacf(beerdiff4, lag.max = 24, main = "")

# DIFFERENCED at lag 4 and lag 1 PACF AND ACF
par(mfrow=c(1,2))
acf(beerdiff4diff1, lag.max=24, main = "")
pacf(beerdiff4diff1, lag.max=24, main = "")


```

Model Selection for SMA(1) 
```{r}

library(qpcR)
AICc<-numeric()
for (p in 0:5) {
  for (q in 0:5) {
    AICc<-round(c(AICc, sarima(beer.tr, p, 1, q, 0, 1, 1, 4, details = FALSE)$AICc),4)
  }
}

AICc<-matrix(AICc, nrow = 6, byrow = TRUE)
rownames(AICc)<-c("p=0","p=1","p=2", "p=3", "p=4", "p=5")
colnames(AICc)<-c("q=0","q=1", "q=2", "q=3", "q=4", "q=5")
AICc
AICc<-data.frame(AICc)
aicc<-setNames(AICc, c("q=0","q=1", "q=2", "q=3", "q=4", "q=5"))
aicc
grid.table(aicc)

```



```{r}

BIC<-numeric()
for (p in 0:5) {
  for (q in 0:5) {
    BIC<- round(c(BIC, sarima(beer.tr, p, 1, q, 0, 1, 1, 4, details = FALSE)$BIC),4)
  }
}

BIC<-matrix(BIC, nrow = 6, byrow = TRUE)
rownames(BIC)<-c("p=0","p=1","p=2", "p=3", "p=4", "p=5")
colnames(BIC)<-c("q=0","q=1","q=2", "q=3", "q=4", "q=5")
BIC
bic<-data.frame(BIC)
bic<-setNames(bic, c("q=0","q=1","q=2", "q=3", "q=4", "q=5"))

library(grid)
grid.table(bic)


```
From SMA(1) AIC selection, we select 

Model Estimation for SMA(1) from AIC (0,1,2)X(1,1,0)4
Model Estimation for SMA(1) from BIC (1,1,1)X(1,1,0)4


```{r}

mod1 <- arima(beer.tr, order = c(0,1,2), seasonal = list(order = c(0,1,1), period = 4))

mod2 <- arima(beer.tr, order = c(1,1,1), seasonal = list(order = c(0,1,1), period = 4))

mod1
mod2
```




Checking Roots on both models
```{r}
#Script for roots
plot.roots <- function(ar.roots=NULL, ma.roots=NULL, size=2, angles=FALSE, special=NULL, sqecial=NULL,my.pch=1,first.col="blue",
second.col="red",main=NULL)
{xylims <- c(-size,size)
omegas <- seq(0,2*pi,pi/500)
temp <- exp(complex(real=rep(0,length(omegas)),imag=omegas))
plot(Re(temp),Im(temp),typ="l",xlab="x",ylab="y",xlim=xylims,ylim=xylims,main=main)
abline(v=0,lty="dotted")
abline(h=0,lty="dotted")
if(!is.null(ar.roots))
{
points(Re(1/ar.roots),Im(1/ar.roots),col=first.col,pch=my.pch)
points(Re(ar.roots),Im(ar.roots),col=second.col,pch=my.pch)
}
if(!is.null(ma.roots))
{
points(Re(1/ma.roots),Im(1/ma.roots),pch="*",cex=1.5,col=first.col)
points(Re(ma.roots),Im(ma.roots),pch="*",cex=1.5,col=second.col)
}
if(angles)
{
if(!is.null(ar.roots))
{
abline(a=0,b=Im(ar.roots[1])/Re(ar.roots[1]),lty="dotted")
abline(a=0,b=Im(ar.roots[2])/Re(ar.roots[2]),lty="dotted")
}
if(!is.null(ma.roots))
{
sapply(1:length(ma.roots), function(j) abline(a=0,b=Im(ma.roots[j])/Re(ma.roots[j]),
lty="dotted"))
}
}
if(!is.null(special))
{
lines(Re(special),Im(special),lwd=2)
}
if(!is.null(sqecial))
{
lines(Re(sqecial),Im(sqecial),lwd=2)
}
}

```


```{r}

# Plotting Roots for Both Models
par(mfrow = c(1,3))
plot.roots(NULL, polyroot (c(1,-0.9663,0.4146)) , main ="Model I MA part")
plot.roots(NULL, polyroot (c (1,0.8019)) , main ="Model I SMA part")
par(mfrow = c(1,3))
plot.roots(NULL, polyroot (c(1,-0.4926)) , main =" Model II MA part")
plot.roots(NULL, polyroot (c(1,-0.4229)) , main =" Model II AR part")
plot.roots(NULL, polyroot (c(1, 0.7603)) , main =" Model II SMA part ")
# normality
par(mfrow = c(2,2))

```

Model1:SARIMA(0,1,2)x(0,1,1)4
AICc = -0.9514499 BIC = -1.905785	
Xt = Zt-0.97Zt-1 + 0.41Zt-2 
NABLA(4)NABLA(Yt) = (1+0.97B-0.41B^2-0.80B^4)

Model 2:SARIMA(1,1,1)x(0,1,1)4
AICc = -0.9171915	BIC = -1.871526



Model Diagnostics 1
```{r}

# Residual Plot Model 1
ts.plot(mod1$residuals, main = "Model 1 Residuals")


```


```{r}

par(mfrow=c(1,2),oma=c(0,0,2,0))
op <- par(mfrow=c(2,2))

acf(mod1$residuals, main = "", lag.max = 24)
pacf(mod1$residuals, main = "", lag.max= 24)
# histogram
hist(mod1$residuals, main = "Histogram of Residuals for Model 1")
# qq plot
qqnorm(mod1$residuals, main = "Normal QQ Plot for Model 1")
qqline(mod1$residuals)

title("Model 1 Diagnostics", outer=TRUE)
par(op)


#Normality Tests
shapiro.test(mod1$residuals)
Box.test(mod1$residuals, lag =4, type = "Box-Pierce", fitdf =2)
Box.test(mod1$residuals, lag =4, type = "Ljung-Box", fitdf =2)
```

Model 2 Diagnostics
```{r}

# Residual Plot Model 2
ts.plot(mod1$residuals, main = "Model 2 Residuals")

```


```{r}

#Model 2 Diagnostics
par(mfrow=c(1,2),oma=c(0,0,2,0))
op <- par(mfrow=c(2,2))

acf(mod2$residuals, main = "", lag.max = 24)
pacf(mod2$residuals, main = "", lag.max=24)
# histogram
hist(mod2$residuals, main = "Histogram of Residuals for Model 2")
# qq plot
qqnorm(mod2$residuals, main = "Normal QQ Plot for Model 2")
qqline(mod2$residuals)

title("Model 2 Diagnostics", outer=TRUE)
par(op)


#Normality Tests
shapiro.test(mod2$residuals)
Box.test(mod2$residuals, lag =4, type = "Box-Pierce", fitdf =2)
Box.test(mod2$residuals, lag =4, type = "Ljung-Box", fitdf =2)

```

Forecasting

```{r}

# Predictions on Transformed Data

pred.log <- predict(mod1, n.ahead=8)

upper.log <- pred.log$pred + 2*pred.log$se 
lower.log <- pred.log$pred - 2*pred.log$se

ts.plot(beer.tr, xlim=c(1, length(beer.tr)+8), main = "Forecasting on Transformed Data", ylab= "y^(0.5)")

lines(upper.log, col="blue", lty = "dashed")
lines(lower.log, col="blue", lty = "dashed")

points((length(beer.tr)+1):(length(beer.tr)+8), pred.log$pred, col ="red")

# Add a legend
legend("bottomright", 
  legend = c("Prediction"), 
  col = c("red"), 
  pch = 1, 
  bty = "o", 
  pt.cex = 1, 
  cex = 1, 
  text.col = "black", 
  horiz = F , 
  inset = c(0.1, 0.1))
```


```{r}

# Predictions on Original Data

pred.original <- pred.log$pred^(1/0.5)


upper <-upper.log^(1/0.5)
lower <-lower.log^(1/0.5)

ts.plot(beer.ts, xlim=c(1, length(beer.ts)+8), main = "Forecasting on Original Data", ylab= "Production in Megalitres")

lines(upper, col="blue", lty = "dashed")
lines(lower, col="blue", lty = "dashed")

points((length(beer.ts)+1):(length(beer.ts)+8), pred.original, col ="red")

# Add a legend
legend("bottomright", 
  legend = c("Prediction"), 
  col = c("red"), 
  pch = 1, 
  bty = "o", 
  pt.cex = 1, 
  cex = 1, 
  text.col = "black", 
  horiz = F , 
  inset = c(0.1, 0.1))


```

```{r}

# zOOM Foreast

ts.plot(beerall.ts, xlim = c(length(beerall.ts)-12, length(beerall.ts)), main = "Observed vs Forecasted Values", ylab = "Production in Megalitres")

# Points for Original Data *
points((length(beer.ts)+1):(length(beer.ts)+8), beerall.ts[147:154], col ="blue")

# Points for Forecasted Data Red
points((length(beer.ts)+1):(length(beer.ts)+8), pred.original, col ="red")

lines((length(beer.ts)+1): (length(beer.ts)+8), upper, lty=2, col = "blue")

lines((length(beer.ts)+1): (length(beer.ts)+8), lower, lty=2, col = "blue")

# Add a legend
legend("bottomright", 
  legend = c("Prediction", "Observed"), 
  col = c("red", 
  "blue"), 
  pch = 1, 
  bty = "o", 
  pt.cex = 1, 
  cex = 1, 
  text.col = "black", 
  horiz = F , 
  inset = c(0.1, 0.1))

```



